---
title: "binomial model for nudi reaction"
author: "Bri Over"
date: "`r Sys.Date()`"
output: html_document
---

faster as got bigger but looking at interaction can see that bigger or smaller did not matter 
as temperature increased
only difference was in 18 compared to 12 (nudis at 18 moved slower)


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
setwd("C:/Users/brian/Documents/Bamfield Fall 2024/DS-ADA/DS/nudi-git")
library(here)
library(glmmTMB)
library(tidyverse)
library(DHARMa)
library(ggeffects)
library(performance)
library(car)
library(mgcv)
library(fitdistrplus)
library(goft)
library(gamlss)
```

```{r}
nudi_data <- read.csv("clean_data/combinedraw.csv")%>%
  mutate(temp_grp = as.factor(temp_grp))
nudi_data$trial_type[nudi_data$trial_type=="conspecific "] <- "conspecific"
# Read in the original dataset

# Create a new variable 'expected_direction' based on the trial_type
#nudi_data$expected_direction <- ifelse(nudi_data$trial_type == "conspecific", 1, 0)

# Create a new variable 'move_as_expected' to indicate whether the observed direction matches the expected direction
#nudi_data$move_as_expected <- ifelse(nudi_data$direction_binom == nudi_data$expected_direction, 1, 0)

# View the structure of the updated dataset to confirm that the new variables have been added
str(nudi_data)

```


```{r}
#glm_binom <- glmmTMB(move_as_expected ~ temp_grp * avg_length_m + trial_type, family = binomial(link = "logit"), data = nudi_data)
#simulateResiduals(glm_binom)%>%
#  plot()

glm_binom2 <- glmmTMB(direction_binom ~ temp_grp * avg_length_m + trial_type, family = binomial(link = "logit"), data = nudi_data)
simulateResiduals(glm_binom2)%>%
  plot()

# Extract the coefficients directly from the model summary
binom_table <- as.data.frame(summary(glm_binom2)$coefficients$cond)

summary(glm_binom2)

# Save the table as a CSV file
write.csv(binom_table, "bri_code/binom_table.csv", row.names = TRUE)


```
did not move based on cues, different cues but same directions
look at difference between 12 and 18 movement
proportion moved tws and proportion moved away


```{r}
# Load necessary libraries
library(dplyr)
library(ggplot2)

# Summarize the data: Count nudibranchs that moved towards or away for each combination of trial_type, temp_grp, and direction_binom
nudi_data_summary <- nudi_data %>%
  group_by(trial_type, temp_grp, direction_binom) %>%
  summarise(count = n(), .groups = 'drop')

# Calculate total number of nudibranchs in each trial_type and temp_grp
nudi_data_totals <- nudi_data %>%
  group_by(trial_type, temp_grp) %>%
  summarise(total = n(), .groups = 'drop')

# Join the summarized data with the total counts to calculate proportions
nudi_data_prop <- nudi_data_summary %>%
  left_join(nudi_data_totals, by = c("trial_type", "temp_grp")) %>%
  mutate(proportion = count / total)  # Calculate proportion for each direction

# Ensure 'direction_binom' is a factor with appropriate levels
nudi_data_prop$direction_binom <- factor(nudi_data_prop$direction_binom, levels = c(0, 1))

# Handle missing data for "towards" direction in case there are no nudibranchs for that direction
nudi_data_prop_complete <- nudi_data_prop %>%
  complete(trial_type, temp_grp, direction_binom, fill = list(count = 0))  # Ensures missing rows are added

# Set any NA values in proportion to 0 (this handles cases where count is 0 or missing)
nudi_data_prop_complete$proportion[is.na(nudi_data_prop_complete$proportion)] <- 0

# Get the maximum proportion to set the y-axis limit
y_max <- max(nudi_data_prop_complete$proportion, na.rm = TRUE)

# Create the bar chart with the correct facet and fixed y-axis
plot_rxn_probs <- ggplot(nudi_data_prop_complete, aes(x = factor(temp_grp), y = proportion, fill = factor(direction_binom))) +
  geom_bar(stat = "identity", position = "dodge") +  # dodge bars so they do not overlap
  facet_wrap(~ trial_type, scales = "free_y",  
             labeller = labeller(trial_type = c("conspecific" = "Conspecific Cues", "predator" = "Predator Cues"))) +  # Rename facet labels
  labs(
    title = "",
    x = "Treatment Group (°C)",
    y = "Proportion of Nudibranchs",
    fill = "Direction"
  ) +
  scale_fill_manual(values = c("0" = "lightblue", "1" = "navyblue"), labels = c("Away", "Towards")) +  # Color customization
  scale_y_continuous(limits = c(0, y_max), expand = c(0, 0)) +  # Fixed y-axis scale across both graphs, ensuring bars touch the x-axis
  scale_x_discrete(expand = c(0.2, 0)) +  # Increase space between the categories on the x-axis
  theme_minimal() +
  theme(
    axis.text.y = element_text(vjust = 1),  # Adjust vertical alignment of y-axis text
    panel.grid.major = element_blank(),  # Remove major gridlines
    panel.grid.minor = element_blank(),  # Remove minor gridlines
    axis.line = element_line(color = "black", linewidth = 0.5),  # Add axis lines for both x and y axes
    axis.ticks = element_line(color = "black", linewidth = 0.5),  # Add ticks to both x and y axes
    axis.title.x = element_text(margin = margin(t = 10)),  # Add space between x-axis labels and numbers
    axis.title.y = element_text(margin = margin(r = 10)),  # Add space between y-axis labels and numbers
    strip.text = element_text(size = 12, face = "bold", margin = margin(b = 20))  # Add space between facet labels and graph
  )

plot_rxn_probs

```
```{r}
summary_table <- nudi_data_prop_complete %>%
  dplyr::select(trial_type, temp_grp, direction_binom, proportion) %>%
  arrange(trial_type, temp_grp, direction_binom)

print(summary_table)


```



```{r}
ggsave(plot = plot_rxn_probs, filename = here("bri_code", "plot_rxn_probs.png"), bg = "white", width = 6, height = 4)

```



```{r}
# Summarize the data to calculate the overall proportions of "towards" and "away" for each temp_grp
nudi_data_summary_overall <- nudi_data %>%
  group_by(temp_grp, direction_binom) %>%
  summarise(count = n(), .groups = 'drop')

# Calculate total number of nudibranchs in each temp_grp
nudi_data_totals_overall <- nudi_data %>%
  group_by(temp_grp) %>%
  summarise(total = n(), .groups = 'drop')

# Join the summarized data with the total counts to calculate proportions
nudi_data_prop_overall <- nudi_data_summary_overall %>%
  left_join(nudi_data_totals_overall, by = "temp_grp") %>%
  mutate(proportion = count / total)  # Calculate proportion for each direction

# Ensure 'direction_binom' is a factor with appropriate levels
nudi_data_prop_overall$direction_binom <- factor(nudi_data_prop_overall$direction_binom, levels = c(0, 1))

# Create the plot showing overall movement for each temp_grp
ggplot(nudi_data_prop_overall, aes(x = factor(temp_grp), y = proportion, fill = factor(direction_binom))) +
  geom_bar(stat = "identity", position = "dodge") +  # Bar chart with dodged bars for direction
  labs(
    title = "All Cues",
    x = "Temperature Group (°C)",
    y = "Proportion of Nudibranchs",
    fill = "Direction"
  ) +
  scale_fill_manual(values = c("0" = "lightblue", "1" = "navyblue"), labels = c("Away", "Towards")) +  # Color customization
  scale_y_continuous(limits = c(0, y_max), expand = c(0, 0)) +  # Fixed y-axis scale across both graphs, ensuring bars touch the x-axis
  scale_x_discrete(expand = c(0.2, 0)) +  # Increase space between the categories on the x-axis
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1),  # Rotate x-axis labels for readability
    axis.text.y = element_text(vjust = 1),  # Adjust vertical alignment of y-axis text
    panel.grid.major = element_blank(),  # Remove major gridlines
    panel.grid.minor = element_blank(),  # Remove minor gridlines
    axis.line = element_line(color = "black", size = 0.5),  # Add axis lines for both x and y axes
    axis.ticks = element_line(color = "black", size = 0.5),  # Add ticks to both x and y axes
    axis.title.x = element_text(margin = margin(t = 10)),  # Add space between x-axis labels and numbers
    axis.title.y = element_text(margin = margin(r = 10)),  # Add space between y-axis labels and numbers
    strip.text = element_text(size = 12, face = "bold", margin = margin(b = 20))  # Add space between facet labels and graph
  )

```






glm_bbinom <- glmmTMB(move_as_expected ~ temp_grp + avg_length_m, family = betabinomial(link = "logit"), data = nudi_data)
simulateResiduals(glm_bbinom)%>%
  plot()
#treats temperature and length separately


# Fit Zero-Inflated Binomial Model
glm_zib <- glmmTMB(direction_binom ~ temp_grp + avg_length_m, 
                   family = binomial(link = "logit"), 
                   zi = ~ temp_grp + avg_length_m, 
                   data = nudi_data)

# Check the model summary
summary(glm_zib)


simulateResiduals(glm_zib) %>% 
  plot()

AIC(glm_bbinom, glm_binom, glm_zib)

#only AIC for binom

summary(glm_binom)

#Temperature effect: The p-value for temp_grp is  not significant at the standard 0.05 significance level: means that, based on this model, there is no strong evidence that temperature affects the nudibranchs' response to conspecific or predatory cues.

#Size effect: Similarly, the p-value for avg_length_m is greater than 0.05 suggesting that the size of the nudibranchs does not significantly impact their direction of movement.


#modeling both the main effects of each variable and their interaction: modeling the effect of temperature, the effect of size, and how the effect of temperature changes depending on the size of the nudibranch

# Fit the model with interaction between temperature and body size
glm_binom_int <- glmmTMB(move_as_expected ~ temp_grp * avg_length_m, family = binomial(link = "logit"), data = nudi_data)

simulateResiduals(glm_binom_int)%>%
  plot()

summary(glm_binom_int)


glm_bbinom_int <- glmmTMB(move_as_expected ~ temp_grp * avg_length_m, family = betabinomial(link = "logit"), data = nudi_data)

simulateResiduals(glm_bbinom_int)%>%
  plot()


summary(glm_binom_int)


# Fit Zero-Inflated Binomial Model
glm_zib_int <- glmmTMB(direction_binom ~ temp_grp * avg_length_m, 
                   family = binomial(link = "logit"), 
                   zi = ~ temp_grp + avg_length_m, 
                   data = nudi_data)

summary(glm_zib_int)


simulateResiduals(glm_zib_int) %>% 
  plot()


AIC(glm_bbinom_int, glm_binom_int, glm_zib_int)

#interaction made this worse somehow: AIC for

