---
title: "binomial model for nudi reaction"
author: "Bri Over"
date: "`r Sys.Date()`"
output: html_document
---

faster as got bigger but looking at interaction can see that bigger or smaller did not matter 
as temperature increased
only difference was in 18 compared to 12 (nudis at 18 moved slower)


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
setwd("C:/Users/brian/Documents/Bamfield Fall 2024/DS-ADA/DS/nudi-git")
library(here)
library(glmmTMB)
library(tidyverse)
library(DHARMa)
library(ggeffects)
library(performance)
library(car)
library(mgcv)
library(fitdistrplus)
library(goft)
library(gamlss)
```

```{r}
nudi_data <- read.csv("clean_data/combinedraw.csv")%>%
  mutate(temp_grp = as.factor(temp_grp))
nudi_data$trial_type[nudi_data$trial_type=="conspecific "] <- "conspecific"
# Read in the original dataset

# Create a new variable 'expected_direction' based on the trial_type
#nudi_data$expected_direction <- ifelse(nudi_data$trial_type == "conspecific", 1, 0)

# Create a new variable 'move_as_expected' to indicate whether the observed direction matches the expected direction
#nudi_data$move_as_expected <- ifelse(nudi_data$direction_binom == nudi_data$expected_direction, 1, 0)

# View the structure of the updated dataset to confirm that the new variables have been added
str(nudi_data)

```


```{r}
#glm_binom <- glmmTMB(move_as_expected ~ temp_grp * avg_length_m + trial_type, family = binomial(link = "logit"), data = nudi_data)
#simulateResiduals(glm_binom)%>%
#  plot()

glm_binom2 <- glmmTMB(direction_binom ~ temp_grp * avg_length_m + trial_type, family = binomial(link = "logit"), data = nudi_data)
simulateResiduals(glm_binom2)%>%
  plot()

#summary(glm_binom)
summary(glm_binom2)


```
did not move based on cues, different cues but same directions
look at difference between 12 and 18 movement
proportion moved tws and proportion moved away


glm_bbinom <- glmmTMB(move_as_expected ~ temp_grp + avg_length_m, family = betabinomial(link = "logit"), data = nudi_data)
simulateResiduals(glm_bbinom)%>%
  plot()
#treats temperature and length separately


# Fit Zero-Inflated Binomial Model
glm_zib <- glmmTMB(direction_binom ~ temp_grp + avg_length_m, 
                   family = binomial(link = "logit"), 
                   zi = ~ temp_grp + avg_length_m, 
                   data = nudi_data)

# Check the model summary
summary(glm_zib)


simulateResiduals(glm_zib) %>% 
  plot()

AIC(glm_bbinom, glm_binom, glm_zib)

#only AIC for binom

summary(glm_binom)

#Temperature effect: The p-value for temp_grp is  not significant at the standard 0.05 significance level: means that, based on this model, there is no strong evidence that temperature affects the nudibranchs' response to conspecific or predatory cues.

#Size effect: Similarly, the p-value for avg_length_m is greater than 0.05 suggesting that the size of the nudibranchs does not significantly impact their direction of movement.


#modeling both the main effects of each variable and their interaction: modeling the effect of temperature, the effect of size, and how the effect of temperature changes depending on the size of the nudibranch

# Fit the model with interaction between temperature and body size
glm_binom_int <- glmmTMB(move_as_expected ~ temp_grp * avg_length_m, family = binomial(link = "logit"), data = nudi_data)

simulateResiduals(glm_binom_int)%>%
  plot()

summary(glm_binom_int)


glm_bbinom_int <- glmmTMB(move_as_expected ~ temp_grp * avg_length_m, family = betabinomial(link = "logit"), data = nudi_data)

simulateResiduals(glm_bbinom_int)%>%
  plot()


summary(glm_binom_int)


# Fit Zero-Inflated Binomial Model
glm_zib_int <- glmmTMB(direction_binom ~ temp_grp * avg_length_m, 
                   family = binomial(link = "logit"), 
                   zi = ~ temp_grp + avg_length_m, 
                   data = nudi_data)

summary(glm_zib_int)


simulateResiduals(glm_zib_int) %>% 
  plot()


AIC(glm_bbinom_int, glm_binom_int, glm_zib_int)

#interaction made this worse somehow: AIC for

